---
title: "Project 1.1"
author: "Alexander Kapustin"
date: "2024-03-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(dplyr)
library(caret)
library(glmnet)
```

```{r}
clean_set <- read.csv("/path/")
```

tPrec: This variable represents the amount of precipitation in millimetres (last hour)

atmosPStatn: This variable represents atmospheric pressure at a station in millibars (mb)

rad: This variable represents Solar radiation $\frac{KJ}{m^2}$

airTemp: This variable represents air temperature(instant) in $^\circ C$ *

dpTemp: This variable represents dew point temperature(instant) in $^\circ C$

airHum: This variable represents air humidity(instant) in %

windDir: This variable represents wind direction in radius degrees ($0-360^\circ$)

windSp: This variable represents wind speed in meters per second

lat: This variable represents latitude

long: This variable represents longitude

(*) Variable of interest ($Y$)

```{r}
#View(clean_set)
clean_set <- clean_set[, -6] # removing prevHrMinTemp

# Rows in original ds
num_rows <- nrow(clean_set)
num_rows
```

```{r}
# 10% of the rows in original ds
num_sample <- round(num_rows * 0.1)
num_sample
```

```{r}
set.seed(42) # Nerd alert! "The Hitchhiker's Guide to the Galaxy"

# Sample 10% of the rows at random
clean_sample <- clean_set[sample(num_rows, num_sample, replace = FALSE), ]
#View(clean_sample)
```

```{r}
# Function for split_group train = 70% test = 30%
split_group <- function(group) {
  # Sort vals in ascending order of date_time
  group <- group %>% arrange(date_time)
  # Non-random, ordered train-test split
  train_set <- group %>% slice(1:round(n() * 0.7))
  test_set <- group %>% slice((round(n() * 0.7) + 1):n())
  return(list(train_set, test_set))
}

# Group by lat and long
grouped_data <- clean_sample %>% group_by(lat, long)
grouped_data <- as.data.frame(grouped_data)
train_test_sets <- grouped_data %>% group_modify(~split_group(.x))

# Train and Test sets
train_set <- bind_rows(train_test_sets[[1]])
train_set <- train_set %>% arrange(row_number()) # Reset row numbers
test_set <- bind_rows(train_test_sets[[2]])
test_set <- test_set %>% arrange(row_number())

#View(train_set)
#View(test_set)
```

```{r}
k = 10 # folds

# Function for polynomial regression with CV
polynomial_cv <- function(data, variable, degrees, k) {
  mse_storage <- numeric(length(degrees))
  
  for (i in seq_along(degrees)) {
    formula <- as.formula(paste("airTemp ~ poly(", variable, ",", degrees[i], ")"))
    model <- train(formula, data = data, method = "lm", trControl = trainControl(method = "cv", number = k))
    mse_storage[i] <- mean(model$results$RMSE^2)
    
  }
  
  return(mse_storage)
  
}

# Degrees
degrees <- 1:5

# Optimal degrees
optimal_degrees <- list()

# Variables loop 
for (variable in c("tPrec", "atmosPStatn", "rad", "dpTemp", "airHum", "windDir", "windSp", "lat", "long")) {
  
  # Run it through the function
  mse_storage <- polynomial_cv(train_set, variable, degrees, k)
  
  # Optimal degrees
  degree_optimal <- which.min(mse_storage)
  
  # If the difference in avg MSE values between the optimal degree 
  # and the next degree is less than 10% it is not significant
  if (degree_optimal > 1 && (mse_storage[degree_optimal - 1] - mse_storage[degree_optimal]) / mse_storage[degree_optimal] < 0.1) {
    degree_optimal <- degree_optimal - 1  # Choosing the lower degree to prevent overfitting
    
  }
  
  # Optimal degree for each variable
  optimal_degrees[[variable]] <- degree_optimal
  
}

best_degrees <- list()

for (variable in names(optimal_degrees)) {
  best_degrees[[variable]] <- paste("The best polynomial degree for ", variable, "is: ", optimal_degrees[[variable]])
}

best_degrees
```

```{r}
# Poly Regression Model
poly_final_samp <- lm(airTemp ~ poly(lat, 4) + poly(long, 4) + 
                  poly(tPrec, 3) + poly(atmosPStatn, 4) + poly(windSp, 4) + 
                  poly(windDir, 4) + poly(dpTemp, 4) + poly(airHum, 4) + poly(rad, 4), 
                data = train_set)
summary(poly_final_samp)
```

```{r}
# Predictions for the test set using the poly regression model
test_predictions_samp <- predict(poly_final_samp, newdata = test_set)
# Accuracy of the model on the test set
test_accuracy_samp <- sqrt(mean((test_set$airTemp - test_predictions_samp)^2))
mean_y_test_samp <- mean(test_set$airTemp)
paste("The model's predictions are off by ", (test_accuracy_samp/mean_y_test_samp)*100, "%")
```

```{r}
# MSE
mse_samp <- mean((test_set$airTemp - test_predictions_samp)^2)
# MAE
mae_samp <- mean(abs(test_set$airTemp - test_predictions_samp))

paste("The Mean Squared Error for the model is: ", mse_samp)
paste("The Mean Absolute Error for the model is: ", mae_samp)
```

```{r}
# test set = 30% of the full clean_set
split_group1 <- function(group) {

  group <- group %>% arrange(date_time)
  test_set1 <- group %>% slice((round(n() * 0.7) + 1):n())

  return(test_set1)
}

grouped_data <- clean_set %>% group_by(lat, long)
grouped_data <- as.data.frame(grouped_data)

test_set1 <- grouped_data %>% group_modify(~split_group1(.x))
test_set1 <- test_set1 %>% arrange(row_number())
```

```{r}
# Predictions for the test set using the poly regression model
test_predictions_samp <- predict(poly_final_samp, newdata = test_set1)
# Accuracy of the model on the test set
test_accuracy_samp <- sqrt(mean((test_set1$airTemp - test_predictions_samp)^2))
mean_y_test_samp <- mean(test_set1$airTemp)
paste("The model's predictions are off by ", (test_accuracy_samp/mean_y_test_samp)*100, "%")
```

```{r}
# MSE
mse_samp <- mean((test_set1$airTemp - test_predictions_samp)^2)
# MAE
mae_samp <- mean(abs(test_set1$airTemp - test_predictions_samp))

paste("The Mean Squared Error for the model is: ", mse_samp)
paste("The Mean Absolute Error for the model is: ", mae_samp)
```
